Concurrent Programming in Harmony

### 2021-07-22 book/paper.tex
Latest

Python is object-oriented, but Harmony is not.  This can lead to some
unexpected differences.  For example, consider the following code:

@@

In Python, this creates two variables $x$ and $y$ referring to the same lock.
In Harmony, the two variables will be two different locks.  If you want two
variables referring to the same lock in Harmony, you would write:

\begin{code}
\begin{verbatim}
x = y = malloc(Lock());
\end{verbatim}
\end{code}

@@

Local variables all declared.  They can be declared in
\texttt{def} statements (i.e., arguments),
\texttt{let} statements, and in \texttt{for} loops.
Also, each method has an implicitly declared
\textit{result} variable, which is initialized to \texttt{None}.
Local variables are tightly scoped and cannot be shared
between threads.
While in theory one method can be declared within another,
they cannot share variables either.
All other variables are global and must be initialized
before any threads are spawned.

@@

There are three classes of precedence.
Application has the highest precedence.  So, \texttt{!}\textit{a b} is interpreted as
\texttt{!}(\textit{a b}) and \textit{a b} \texttt{+} \textit{c d} is interpreted as
(\textit{a b}) \texttt{+} (\textit{c d}).
Unary operators have the next highest precedence,
and the remaining operators have the lowest precedence.
For example, $-2 + 3$ evaluates to 1, not $-5$.

@@

Harmony allows pattern matching against nested tuples in various
language constructs.
\index{pattern matching}%
The following are the same in Python and Harmony:
\begin{itemize}
\item \texttt{$x,$ = 1,}: assigns 1 to $x$;
\item \texttt{$x$, $y$ = 1, (2, 3)}: assigns 1 to $x$ and (2, 3) to $y$;
\item \texttt{$x$, ($y$, $z$) = 1, (2, 3)}: assigns 1 to $x$, 2 to $y$, and
3 to $z$;
\item \texttt{$x$, ($y$, $z$) = 1, 2;} generates an runtime error because 2 cannot
be matched with ($y$, $z$);
\item \texttt{$x[0]$, $x[1]$ = $x[1]$, $x[0]$;} swaps the first two elements of list $x$.
\end{itemize}

@@

Pattern matching can also be used in \texttt{const} and \texttt{let}
statements.

@@

In case of dictionaries, you can access both keys and values:
\begin{code}
\texttt{for} $k:v$ in \{ .a: 1, .b: 2 \}: ...
\end{code}

@@

@@

Finally, I would like to thank my family who had to suffer as I obsessed
over writing the code and the book during the turbulent months of
May and June 2020.

>>>>

Python is object-oriented, but Harmony is not.  For Python programmers,
this can lead to some unexpected differences.
For example, consider the following code:

@@

In Python, this creates two variables $x$ and $y$ referring to the same lock.
In Harmony, the two variables will be two different locks.  If you want two
variables referring to the same lock in Harmony, you might write:

\begin{code}
\begin{verbatim}
lock = Lock()
x = y = ?lock
\end{verbatim}
\end{code}

@@

Local variables all declared.  They can be declared in
\texttt{def} statements (i.e., arguments),
\texttt{let} statements, and in \texttt{for} loops.
Also, each method has an implicitly declared
\textit{result} variable, which is initialized to \texttt{None}.
Local variables are tightly scoped and cannot be shared
between threads.
While in theory one method can be declared within another,
they cannot share local variables either.
All other variables are global and must be initialized
before any threads are spawned.

@@

There are three levels of precedence.
Application has the highest precedence.  So, \texttt{!}\textit{a b} is interpreted as
\texttt{!}(\textit{a b}) and \textit{a b} \texttt{+} \textit{c d} is interpreted as
(\textit{a b}) \texttt{+} (\textit{c d}).
Unary operators have the next highest precedence,
and the remaining operators have the lowest precedence.
For example, $-2 + 3$ evaluates to 1, not $-5$.

@@

Harmony allows pattern matching against nested tuples in various
language constructs.
\index{pattern matching}%
The following are the same in Python and Harmony:
\begin{itemize}
\item \texttt{$x,$ = 1,}: assigns 1 to $x$;
\item \texttt{$x$, $y$ = 1, (2, 3)}: assigns 1 to $x$ and (2, 3) to $y$;
\item \texttt{$x$, ($y$, $z$) = 1, (2, 3)}: assigns 1 to $x$, 2 to $y$, and
3 to $z$;
\item \texttt{$x$, ($y$, $z$) = 1, 2}: generates an runtime error because 2 cannot
be matched with ($y$, $z$);
\item \texttt{$x[0]$, $x[1]$ = $x[1]$, $x[0]$}: swaps the first two elements of list $x$.
\end{itemize}

@@

Pattern matching can also be used in \texttt{const}, \texttt{let},
and \texttt{select} statements.

@@

In case of dictionaries, you can access both keys and values (not
supported in Python):
\begin{code}
\texttt{for} $k:v$ in \{ .a: 1, .b: 2 \}: ...
\end{code}

@@

Kevin Sun and Anthony Yang built a beautiful VSCode extension for Harmony
called HarmonyLang and proceeded to build an animator for Harmony
executions and two cloud-based Harmony offerings, which you can can learn
about at \url{http://harmony.cs.cornell.edu}.  They also developed much of
that web site.  Later they were joined by Shi Chong Zhao, who also
made significant contributions.

@@

Finally, I would like to thank my family who had to suffer as I obsessed
over writing the code and the book, at home, during the turbulent months of
May and June 2020.
